---
# dont run the code 
eval: false
echo: false
---

# Energy-Intensive Industries Indices

This Code focuses on the production indices for the **five most energy-intensive industries** in the European Union, weighted by their gross value added.

## Imports

```{r}
library(tidyverse)  # Data manipulation
library(eurostat)   # Access Eurostat data
library(wespggplot2) # Custom plotting theme
```

## Load Data

We load the official data from Eurostat using the API (saved as RDS files for reproducibility):

```{r}
# Check access to Eurostat API
check_access_to_data()
```

```{r}
# Uncomment the following to directly pull data from Eurostat (commented out for reproducibility):
df_prod <- get_eurostat("sts_inpr_q", filter = list(geo="EU27_2020"))
saveRDS(df_prod, "df_prod.rds")

# Load saved RDS files (for reproducibility)
df_prod <- readRDS("df_prod.rds")
```

## Energy-Intensive Industries Data

### Weighting Data (Gross Value Added by Sector)

We focus on the **five most energy-intensive sectors** in manufacturing:

* **C17**: Manufacture of Paper
* **C18**: Printing and reproduction of recorded media
* **C20**: Manufacture of chemicals and pharmaceuticals
* **C23**: Manufacture of other non-metallic mineral products
* **C24**: Manufacture of basic metals

The data for these sectors is weighted according to their share in gross value added. We also estimate the value for **C20** (Manufacture of chemicals) from the Draghi Report due to confidentiality in Eurostat's dataset.

```{r}
df_weights <- get_eurostat("nama_10_a64", filter = list(geo="EU27_2020", time="2021"))
saveRDS(df_weights, "df_weights.rds")
```


```{r}
df_weights <- readRDS("df_weights.rds")
```

```{r}
# Adjust weights for confidentiality (C20 is estimated from Draghi Report)
weights <- df_weights %>%
  filter(
    nace_r2 %in% c("C17", "C18", "C20", "C23", "C24"),
    na_item == "B1G",       # Gross Value Added
    unit == "CP_MEUR"       # Million Euros
  ) %>%
  mutate(
    values = case_when(
      nace_r2 == "C20" ~ 144000, # Value from Draghi Report for C20
      TRUE ~ values
    ),
    weight = values / sum(values) # Normalize the weights
  ) %>%
  select(nace_r2, weight)
```

### Composite Index for Energy-Intensive Industries

We now calculate a composite index for the **energy-intensive industries**, weighting each sector’s production index by its share in gross value added.

```{r}
df_prod_cen <- df_prod %>%
  filter(
    time >= "2018-01-01",
    nace_r2 %in% c("C17", "C18", "C20", "C23", "C24"),
    s_adj == "SCA",       # Seasonally adjusted data
    unit == "I21"         # Index with base 2021 = 100
  ) %>%
  left_join(weights, by = "nace_r2") %>%
  group_by(time) %>%
  mutate(C_EN = sum(values * weight)) %>%  # Weighted sum for composite index
  ungroup() %>%
  select(time, C_EN) %>%
  reframe(
    time = time,
    values = C_EN,
    nace_r2 = "C_EN"       # Label for composite index
  ) %>%
  distinct()
```

## Rebase Data to 2021-01-01 = 100

We rebase the composite index so that January 2018 is set to 100 for consistency.

```{r}
df_prod_cen <- df_prod_cen %>%
  group_by(nace_r2) %>%
  mutate(
    base_value = values[time == "2018-01-01"],
    values = (values / base_value) * 100
  ) %>%
  select(-base_value) %>%
  ungroup()

# saveRDS(df_prod_cen, "df_energy_intensive_indices.rds")
```

## Plot

We now create a plot for the composite index of **energy-intensive industries**, using a custom theme.

```{r}
p_indices <- df_prod_cen %>%
  ggplot(aes(x = time, y = values, color = nace_r2)) +
  geom_line() +
  geom_hline(yintercept = 100, linetype = "dotted", color = "#7F7F7F") +
  scale_color_manual(
    labels = c(
      "C_EN" = "Energy-intensive Industries"
    ),
    values = c(
      "C_EN" = "#009EDB"
    )
  ) +
  labs(
    subtitle = "Trends in Energy-Intensive Industries in the European Union",
    title = "Figure X: Energy-Intensive Industries Production Index",
    x = "Time",
    color = "Industry",
    tag = "                       Jan 2018 = 100",
    caption = "Source: UN DESA, based on Eurostat. Notes: Production index for energy-intensive industries is weighted by gross value added for the five most energy-intensive manufacturing sectors (NACE Classification: C17, C18, C20, C23, C24)."
  ) +
  wesp_theme() +
  scale_x_date(date_breaks="1 year", date_labels="%Y")

p_indices
```

## Save Index as CSV

Finally, we save the composite index data as a CSV file for further analysis.

```{r}
write_csv(df_prod_cen, "../EII_index.csv")
```


